---
title: "MaetzigEtAl2017 ICCM summary/analysis for SR/OR with reflexives"
author: "Paul Mätzig"
date: "5/11/2017"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height=8,fig.width=8)
library(dplyr)
library(ggplot2)
library(plotly)
library(plot3D)         ## for 3D scatterplots
library(scatterplot3d)  ## for 3D scatterplots
library(ggplot2)
library(ggdendro)
library(factoextra)
#opts_chunk$set(warning=FALSE,comment=NA,fig.height=8,fig.width=8)
```

```{r loaddata,echo=FALSE}
caplan_data <- read.csv2('../data/CaplanEtAl_2015_reflexives.csv', header=TRUE, stringsAsFactors = FALSE) 
caplan_data$sent <- as.character(caplan_data$sent)
caplan_data$acc <- as.numeric(caplan_data$acc)
caplan_data_SR <- filter(caplan_data, sent=='SSREF')
caplan_data_OR <- filter(caplan_data, sent=='SOREF')

model_data <- read.csv('../data/response_accuracies_CAPLAN15_SOREFL_20170526_1533.csv', header=TRUE)
model_data_SR <- filter(model_data, condition=='SR-REFL')
model_data_OR <- filter(model_data, condition=='OR-REFL')
```

```{r preprocessdata_caplan,echo=FALSE}
caplan_data_SR_acc <- summarise(group_by(caplan_data_SR, subj,item), sum(acc, na.rm=TRUE) / n())
#head(caplan_data_SR_acc)
#summary(caplan_data_SR_acc)
#xtabs(~subj+item, caplan_data_SR_acc)
colnames(caplan_data_SR_acc) <- c('subj', 'item', 'acc')
caplan_data_SR_acc <- summarise(group_by(caplan_data_SR_acc, subj), sum(acc) / n())
colnames(caplan_data_SR_acc)<-c('subj', 'acc')
#summary(caplan_data_SR_acc)

caplan_data_OR_acc <- summarise(group_by(caplan_data_OR, subj,item), sum(acc, na.rm=TRUE) / n())
colnames(caplan_data_OR_acc) <- c('subj', 'item', 'acc')
#head(caplan_data_OR_acc)
#summary(caplan_data_OR_acc)
#xtabs(~subj+item, caplan_data_OR_acc)
caplan_data_OR_acc <- summarise(group_by(caplan_data_OR_acc, subj), sum(acc) / n())
colnames(caplan_data_OR_acc)<-c('subj', 'acc')
#summary(caplan_data_OR_acc)

slots_SR <- matrix(rep(NA, dim(model_data_SR)[1]*dim(caplan_data_SR_acc)[1]), ncol=dim(caplan_data_SR_acc)[1])
slots_OR <- matrix(rep(NA, dim(model_data_OR)[1]*dim(caplan_data_OR_acc)[1]), ncol=dim(caplan_data_OR_acc)[1])

for (i in 1:length(caplan_data_SR_acc$subj)) {
  tmp_acc <- caplan_data_SR_acc$acc[i]
  slots_SR[,i] <- abs(model_data_SR$acc_embV - tmp_acc)
}

for (i in 1:length(caplan_data_OR_acc$subj)) {
  tmp_acc <- caplan_data_OR_acc$acc[i]
  slots_OR[,i] <- abs(model_data_OR$acc_embV - tmp_acc)
}

colnames(slots_SR) <- paste("s", caplan_data_SR_acc$subj, sep = "")
colnames(slots_OR) <- paste("s", caplan_data_OR_acc$subj, sep = "")

model_data_SR <- cbind(model_data_SR, slots_SR)
model_data_OR <- cbind(model_data_OR, slots_OR)

get_min_params <- function(data, subject) {
  ref <- which(data[c(subject)]==min(data[c(subject)]))
  params <- data[ref, c("GA", "DAT", "ANS")]
  return(params)
}

subjects_SR <- colnames(model_data_SR)[12:dim(model_data_SR)[2]]
subjects_OR <- colnames(model_data_OR)[12:dim(model_data_OR)[2]]

results_SR <- data.frame()
for (i in 1:length(subjects_SR)) {
  tmp <- get_min_params(model_data_SR, subjects_SR[i])
  tmp$subj <- rep(subjects_SR[i], dim(tmp)[1])
  results_SR <- rbind(results_SR, tmp, make.row.names=FALSE)
}

#subjects<-unique(results_SR$subj)
#for(i in 1:length(subjects)){
#print(dim(subset(results_SR,subj==subjects[i]))[1])
#}
  
results_OR <- data.frame()
for (i in 1:length(subjects_OR)) {
  tmp <- get_min_params(model_data_OR, subjects_OR[i])
  tmp$subj <- rep(subjects_OR[i], dim(tmp)[1])
  results_OR <- rbind(results_OR, tmp, make.row.names=FALSE)
}

## splitting data by SR/OR condition and by control/IWA group
results_SR_controls <- results_SR[1:(min(which(results_SR$subj=="s54001"))-1), ]
results_SR_iwa <- results_SR[min(which(results_SR$subj=="s54001")):dim(results_SR)[1], ]
results_OR_controls <- results_OR[1:(min(which(results_OR$subj=="s54001"))-1), ]
results_OR_iwa <- results_OR[min(which(results_OR$subj=="s54001")):dim(results_OR)[1], ]
```

```{r, echo=FALSE}
if(1){
## averaging values for each subject:
control_subj<-unique(results_SR_controls$subj)
results_SR_c_mn <- data.frame()
for(i in control_subj){
      tmp<-subset(results_SR_controls,subj==i)
      temp<-data.frame(GA=mean(tmp$GA),DAT=mean(tmp$DAT),ANS=mean(tmp$ANS),subj=i)
      results_SR_c_mn<-rbind(results_SR_c_mn,
                        temp,make.row.names=FALSE)
}

control_subj<-unique(results_OR_controls$subj)
results_OR_c_mn <- data.frame()
for(i in control_subj){
      tmp<-subset(results_OR_controls,subj==i)
      temp<-data.frame(GA=mean(tmp$GA),DAT=mean(tmp$DAT),ANS=mean(tmp$ANS),subj=i)
      results_OR_c_mn<-rbind(results_OR_c_mn,
                        temp,make.row.names=FALSE)
}

iwa_subj<-unique(results_SR_iwa$subj)
results_SR_i_mn <- data.frame()
for(i in iwa_subj){
      tmp<-subset(results_SR_iwa,subj==i)
      temp<-data.frame(GA=mean(tmp$GA),DAT=mean(tmp$DAT),ANS=mean(tmp$ANS),subj=i)
      results_SR_i_mn<-rbind(results_SR_i_mn,
                        temp,make.row.names=FALSE)
}

iwa_subj<-unique(results_OR_iwa$subj)
results_OR_i_mn <- data.frame()
for(i in iwa_subj){
      tmp<-subset(results_OR_iwa,subj==i)
      temp<-data.frame(GA=mean(tmp$GA),DAT=mean(tmp$DAT),ANS=mean(tmp$ANS),subj=i)
      results_OR_i_mn<-rbind(results_OR_i_mn,
                        temp,make.row.names=FALSE)
}

results_SR_controls<-results_SR_c_mn
results_OR_controls<-results_OR_c_mn
results_SR_iwa<-results_SR_i_mn
results_OR_iwa<-results_OR_i_mn
}

# coding subjects as factors
results_SR_controls$subj <- as.integer(factor(results_SR_controls$subj))
results_OR_controls$subj <- as.integer(factor(results_OR_controls$subj))

results_SR_iwa$subj <- as.integer(factor(results_SR_iwa$subj))
results_OR_iwa$subj <- as.integer(factor(results_OR_iwa$subj))
```

```{r}
model_data_refl_acc_SR <- subset(model_data, model_data$condition == "SR-REFL")
model_data_refl_acc_OR <- subset(model_data, model_data$condition == "OR-REFL")

head(model_data_refl_acc_OR)
head(model_data_refl_acc_SR)

plot(model_data_refl_acc_SR$acc_refl ~ model_data_refl_acc_SR$simset, main="SR-REFL")
plot(model_data_refl_acc_OR$acc_refl ~ model_data_refl_acc_OR$simset, main="OR-REFL")

plot(model_data_refl_acc_SR$GA, model_data_refl_acc_SR$acc_refl, main="GA, SR-REFL")
plot(model_data_refl_acc_OR$GA, model_data_refl_acc_OR$acc_refl, main="GA, OR-REFL")

plot(model_data_refl_acc_SR$DAT, model_data_refl_acc_SR$acc_refl, main="DAT, SR-REFL")
plot(model_data_refl_acc_OR$DAT, model_data_refl_acc_OR$acc_refl, main="DAT, OR-REFL")

plot(model_data_refl_acc_SR$ANS, model_data_refl_acc_SR$acc_refl, main="ANS, SR-REFL")
plot(model_data_refl_acc_OR$ANS, model_data_refl_acc_OR$acc_refl, main="ANS, OR-REFL")

```

```{r plot3d_1,echo=FALSE, eval=TRUE}
#pdf(file="../manuscript/figures/SR_controls.pdf", paper="a4")
#par(mfrow=c(2,2))
scatter3D(jitter(results_SR_controls$GA,factor=4),
          jitter(results_SR_controls$DAT,factor=4),
          jitter(results_SR_controls$ANS), colvar=NULL,
          col='black', pch = 19, cex = 1, phi=0,
          xlab="GA", ylab="DAT", zlab="ANS", ticktype="detailed", bty="g",main="Controls, SR")

scatter3D(jitter(results_SR_iwa$GA,factor=4),
          jitter(results_SR_iwa$DAT,factor=4),
          jitter(results_SR_iwa$ANS,factor=4),
          colvar=NULL, col='black', pch = 19, cex = 1, phi=0,
          xlab="GA", ylab="DAT", zlab="ANS", ticktype="detailed", bty="g",main="IWA, SR")

scatter3D(jitter(results_OR_controls$GA,factor=4),
          jitter(results_OR_controls$DAT,factor=4),
          jitter(results_OR_controls$ANS,factor=4),
          colvar=NULL, col='black', pch = 19, cex = 1, phi=0,
          xlab="GA", ylab="DAT", zlab="ANS", ticktype="detailed", bty="g",main="Controls, OR")

scatter3D(jitter(results_OR_iwa$GA,factor=4),
          jitter(results_OR_iwa$DAT,factor=4),
          jitter(results_OR_iwa$ANS,factor=4),
          colvar=NULL, col='black', pch = 19, cex = 1, phi=0,
          xlab="GA", ylab="DAT", zlab="ANS", ticktype="detailed", bty="g",main="IWA, OR")
```

```{r plot3d_2a,echo=FALSE,eval=FALSE}
# base R plots without colours and jitter (can be added at will)
# plot(results_SR_iwa$GA, results_SR_iwa$DAT, pch=19, xlab="GA", ylab="DAT")
# plot(results_SR_iwa$GA, results_SR_iwa$ANS, pch=19, xlab="GA", ylab="ANS")
# plot(results_SR_iwa$DAT, results_SR_iwa$ANS, pch=19, xlab="DAT", ylab="ANS")
print(p_SR_control_GA_DAT <- ggplot(data=results_SR_iwa, aes(x=GA, y=DAT)) + geom_jitter(aes(colour=subj)) + ggtitle("SR iwa, GA and DAT"))
print(p_SR_control_GA_ANS <- ggplot(data=results_SR_iwa, aes(x=GA, y=ANS)) + geom_jitter(aes(colour=subj)) + ggtitle("SR iwa, GA and ANS"))
print(p_SR_control_DAT_ANS <- ggplot(data=results_SR_iwa, aes(x=DAT, y=ANS)) + geom_jitter(aes(colour=subj)) + ggtitle("SR iwa, DAT and ANS"))
#dev.off()
```

```{r plot1d_1,echo=FALSE}
## comparing distributions:
op<-par(
    mar=c(2,4.5,4,2), 
    mfrow=c(1,3),pty="s",
    oma = c(2,10,2,2)
    )
plot(density(results_SR_controls$GA),main="Subject relatives",xlim=c(.2,1.1),ylab="Density",
     xlab="goal activation \n (resource reduction)")
lines(density(results_SR_iwa$GA),lty=2)
legend(x=.3,y=2,legend=c("controls","IWA"),lty=1:2)

plot(density(results_SR_controls$DAT),xlab="Default action time\n (slowed processing)",main="Subject relatives",xlim=c(.05,.1),ylim=c(0,30))
lines(density(results_SR_iwa$DAT),xlab="DAT",lty=2)

plot(density(results_SR_controls$ANS),xlab="Noise (intermittent deficiency)",main="Subject relatives",xlim=c(.15,.45))
lines(density(results_SR_iwa$ANS),lty=2)
```

```{r}
op<-par(
    mar=c(2,4.5,4,2), 
    mfrow=c(1,3),pty="s",
    oma = c(2,10,2,2)
    )

#plot(density(results_SR_iwa$GA),xlab="GA",main="IWA, SR")
#plot(density(results_SR_iwa$DAT),xlab="DAT",main="IWA, SR")
#plot(density(results_SR_iwa$ANS),xlab="ANS",main="IWA, SR")

#op<-par(mfrow=c(2,3),pty="s")
plot(density(results_OR_controls$GA),xlab="goal activation \n (resource reduction)",main="Object relatives",xlim=c(.2,1.1),ylim=c(0,3))
lines(density(results_OR_iwa$GA),xlab="GA",lty=2)
legend(x=.3,y=3,legend=c("controls","patients"),
       lty=1:2)

plot(density(results_OR_controls$DAT),xlab="Default action time \n (slowed processing)",main="Object relatives",xlim=c(.05,.1))
lines(density(results_OR_iwa$DAT),xlab="DAT",lty=2)

plot(density(results_OR_controls$ANS),xlab="Noise (intermittent deficiency)",main="Object relatives",xlim=c(.15,.45),ylim=c(0,7))
lines(density(results_OR_iwa$ANS),xlab="ANS",lty=2)

#plot(density(results_OR_iwa$GA),xlab="GA",main="IWA, OR")
#plot(density(results_OR_iwa$DAT),xlab="DAT",main="IWA, OR")
#plot(density(results_OR_iwa$ANS),xlab="ANS",main="IWA, OR")
```

# Hierarchical clustering

[pmaetzig, 2017-05-28] I just copied the steps from the previous ICCM analysis file for now, may need to change methods
in hierarchical clustering.

I don't see much of a difference between the two groups in the clustering.

In SRs I chose centroid as the method because that gives the best discrimination ability between controls and iwa.

```{r}
results_SR_controls$type <- "control"
results_SR_iwa$type <- "iwa"
results_OR_controls$type <- "control"
results_OR_iwa$type <- "iwa"
```

```{r controlsSR, echo=FALSE}
clustersSRc<-hclust(dist(results_SR_controls[,1:3],
                         method="euclidean"),method="centroid")

hcdata2 <- dendro_data(clustersSRc, 
                       type="rectangle")


ggdendrogram(hcdata2,rotate=TRUE, size=2) + 
  ggtitle("Hierarchical clustering, SR controls")+  coord_flip()
```

```{r iwaSR, echo=FALSE}
clustersSRi<-hclust(dist(results_SR_iwa[,1:3],
                         method="euclidean"),method="centroid")

hcdata2 <- dendro_data(clustersSRi, 
                       type="rectangle")

ggdendrogram(hcdata2,rotate=TRUE, size=2) + 
  ggtitle("Hierarchical clustering, SR iwa")+  coord_flip()
```


For ORs I chose complete as method because it gives the best discriminative performance between controls and iwa.

```{r controlsOR, echo=FALSE}
clustersORc<-hclust(dist(results_OR_controls[,1:3],
                         method="euclidean"),method="complete")

hcdata2 <- dendro_data(clustersORc, 
                       type="rectangle")

ggdendrogram(hcdata2,rotate=TRUE, size=2) + 
  ggtitle("Hierarchical clustering, OR controls")+  coord_flip()
```

```{r iwaOR, echo=FALSE}
clustersORi<-hclust(dist(results_OR_iwa[,1:3],
                         method="euclidean"),method="complete")

hcdata2 <- dendro_data(clustersORi, 
                       type="rectangle")

ggdendrogram(hcdata2,rotate=TRUE, size=2) + 
  ggtitle("Hierarchical clustering, OR iwa")+  coord_flip()
```

# Discrimination ability using hierarchical clustering

The discrimination ability is OK for controls, but below 50\% for iwa, suggesting that the three parameters don't give us much ability to discriminate controls and iwa.

```{r}
results_OR<-rbind(results_OR_controls,results_OR_iwa)
results_SR<-rbind(results_SR_controls,results_SR_iwa)
```

```{r}
results_SR$subject_full <- paste(results_SR$type, results_SR$subj, sep="-")
results_SR$subject_full <- as.character(results_SR$subject_full)
clusters_SR <- hclust(dist(results_SR[, 1:3]),method="centroid")
plot(clusters_SR, labels=results_SR$subject_full)
clusterCut_SR <- cutree(clusters_SR, 2)
## not able to identify aphasics well
cluster_table_SR <- table(clusterCut_SR, results_SR$type)
print(cluster_table_SR)
# SR, controls
(cluster_table_SR[1,1]/(cluster_table_SR[1,1] + cluster_table_SR[2,1]))
# SR, iwa
(cluster_table_SR[2,2]/(cluster_table_SR[1,2] + cluster_table_SR[2,2]))
```


```{r}
results_OR$subject_full <- paste(results_OR$type, results_OR$subj, sep="-")
results_OR$subject_full <- as.character(results_OR$subject_full)
clusters_OR <- hclust(dist(results_OR[, 1:3]),method="centroid")
plot(clusters_OR, labels=results_OR$subject_full)
clusterCut_OR <- cutree(clusters_OR, 2)
## so-so discrimination ability for aphasics:
cluster_table_OR <- table(clusterCut_OR, results_OR$type)
print(cluster_table_OR)
# OR, controls
(cluster_table_OR[1,1]/(cluster_table_OR[1,1] + cluster_table_OR[2,1]))
# OR, iwa
(cluster_table_OR[2,2]/(cluster_table_OR[1,2] + cluster_table_OR[2,2]))
```

# Discrimination ability using k-means

I'm using the `factoextra` package for an automatised workflow for k-means and their implementation of the gap statistic (cf. Hastie, Tibshirani \& Friedman, 2008 2nd ed, p. 519f.). Using the `eclust` function with method k-means and a k equal to NULL results in estimation of optimal cluster size using the gap statistic.

CODE CHUNKS TAKEN OUT BECAUSE NOT NEEDED

```{r, echo=FALSE, eval=FALSE}
library(factoextra)
SR_clusters <- eclust(results_SR[, 1:3], k=2)
kmeans_table_SR <- table(SR_clusters$cluster, results_SR$type)
# SR, controls
(kmeans_table_SR[1,1]/(kmeans_table_SR[1,1] + kmeans_table_SR[2,1]))
# SR, iwa
(kmeans_table_SR[2,2]/(kmeans_table_SR[1,2] + kmeans_table_SR[2,2]))
```

```{r, echo=FALSE, eval=FALSE}
OR_clusters <- eclust(results_OR[, 1:3], k=2)
kmeans_table_OR <- table(OR_clusters$cluster, results_OR$type)
# OR, controls
(kmeans_table_OR[1,1]/(kmeans_table_OR[1,1] + kmeans_table_OR[2,1]))
# OR, iwa
(kmeans_table_OR[2,2]/(kmeans_table_OR[1,2] + kmeans_table_OR[2,2]))
```

# Distributions of subjects

```{r}
results_SR$RCType<-"SR"
results_OR$RCType<-"OR"
results_all<-rbind(results_SR,results_OR)

results_all$DATnormal<-ifelse(results_all$DAT > .06,
       "elevated","normal")
xtabs(~DATnormal+type+RCType,results_all)

results_all$ANSnormal<-ifelse(results_all$ANS > .2,
       "elevated","normal")
xtabs(~ANSnormal+type+RCType,results_all)

results_all$GAnormal<-ifelse(results_all$GA < .8,
       "elevated","normal")
xtabs(~GAnormal+type+RCType,results_all)

## two impairments:
results_all$DATGAnormal<-ifelse(results_all$DAT > .06 & results_all$GA < .8,
       "elevated","normal")
xtabs(~DATGAnormal+type+RCType,results_all)

results_all$ANSGAnormal<-ifelse(results_all$ANS > .2 & results_all$GA < .8,
       "elevated","normal")
xtabs(~ANSGAnormal+type+RCType,results_all)

results_all$DATANSnormal<-ifelse(results_all$DAT > .06 & results_all$ANS > .2,
       "elevated","normal")
xtabs(~DATANSnormal+type+RCType,results_all)

## all three
results_all$GADATANSnormal<-ifelse(results_all$GA < 0.8 & results_all$DAT > .06 & results_all$ANS > .2,
       "elevated","normal")
xtabs(~GADATANSnormal+type+RCType,results_all)
```

