---
title: "Visualising the effect of a low W parameter setting"
author: "Paul Mätzig"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demo paramsearch analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

This is a demo of the core functions of this package, as part of my master's
thesis. I will show a minimal example of an analysis of a `run-param-search-em`
simulation of the ACT-R model (cf. Lewis & Vasishth, 2005, Engelmann et al., 
2013) -- for the standard setting and a low value of the $W$ parameter. This 
is intended to model low working memory capacity, with the outlook of modelling
aphasic sentence comprehension.

## Workflow

Because the output format of the fixations that the ACT-R model produces is
fairly uniform across different sentences and experiments, this package
exploits the `paramsearch.txt` file to automatically extract regions of
interest and assigning paramsearch sets and parameters settings as names of the
data frames.

The way this package is intended to be used for analysis of paramsearch
analyses is to either

1) copy this vignette as an rmarkdown file into the directory within
`paramsearch/` and follow the workflow given here, or
2) in this same directory, write a short script yourself that calls the
functions, using this workflow as a guide.

Here, we're going to concentrate on visualising the effect of a low $W$ parameter
setting on total fixation time (TFT). A look at `paramsearch.txt` shows that,
additionally, a simulation with a high $W$ has been run, but we're going to
ignore this.

```{r}
library(ACTRPredictions2016)
library(dplyr)
library(ggplot2)
p <- get_parameters("../data/paramsearch.txt")
head(p)
```

The `paramsearch.txt` file contains metadata on the simulation that was run with
the ACT-R model. We can automatically read and correctly name all the fixation
files in the directory specified by `path`, via the function `get_data`. The
result is a list of data frames.

```{r}
fixation_list <- get_data("../data", p)
summary(fixation_list)
```

To calculate reading measures from the raw fixation data, we will use `em2`
which is wrapped in the function `calculate_reading_measures`:

```{r}
data_list <- calculate_reading_measures(fixation_list)
head(data_list[[1]])
```

We will only look at the main and embedded verb regions, to visualise an effect.
For now, there is the function `extract_main_emb_V` (this will be generalised
in a later version of this package):

```{r}
data_list <- extract_main_emb_V(data_list)
head(data_list[[1]])
```

Calculate means and standard error of the mean by region and condition:

```{r}
## This is how I did it before:
# tapply(data_list[[1]]$TFT, INDEX=list(data_list[[1]]$condition, data_list[[1]]$roi), mean, na.rm=TRUE)
# tapply(data_list[[1]]$TFT, INDEX=list(data_list[[1]]$condition, data_list[[1]]$roi), standard_error)

# And here's how to do it with this package:
results_list <- mean_se_by_region(data_list, measure="TFT")
print(head(results_list[[1]]))
print(results_list[[1]]$`mean(TFT)`)
print(results_list[[1]]$`standard_error(TFT)`)
```

Quick test if plotting with `ggplot2::qplot` works:

```{r}
## Multiple plots next to each other not working yet. How to do that?
## (for loop doesn't get evaluated in a chunk
# par(mfrow=(c(1, 2)))
plot_by_region(results_list[[2]], maintitle="Standard W parameter setting", legendpos=c(0, 0.2))
plot_by_region(results_list[[1]], maintitle="Low W parameter setting", legendpos=c(0, 0.2))
```
