---
title: "Data analysis of Wu et al 2016"
author: "Shravan Vasishth"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data analysis of Wu et al 2016}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r loadlibraries, echo=FALSE, include=FALSE}
library(lme4)
#library(RePsychLing)
#library(car)
library(reshape)
library(ggplot2)
library(plyr)
library(grid)
library(lattice)
library(xtable)
library(MASS)

source("../R/multiplot.R")
source("../R/plotresults.R")
source("../R/coeflmer.R")
source("../R/regionmeans.R")
source("../R/regionplot.R")
source("../R/magnifytext.R")
```

```{r, echo = FALSE}
      knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


# Experiment 1

```{r preprocesse1e2,echo=FALSE,include=FALSE,eval=FALSE}
e1<-read.table("../data/Expt1.txt",header=TRUE)

## create region table:
## recode regions of interest:
region<-c(0:6,0:7,0:7,0:8)
length(region)
cond<-rep(letters[1:4],c(7,8,8,9))
condaroi<-c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO")
condbroi<-c("CL","RCSubj","RCV","de","headnoun","Adv","MainV","MainO")
condcroi<-c("BEI","RCSubj","RCV","de","headnoun","Adv","MainV","MainO")
conddroi<-c("CL","BEI","RCSubj","RCV","de","headnoun","Adv","MainV","MainO")

condroi<-c(condaroi,condbroi,condcroi,conddroi)
regiontable<-data.frame(cond=cond,reg=region,roi=condroi)

e1temp<-merge(e1,regiontable,by.x=c("condition","region"),by.y=c("cond","reg"))

e1<-e1temp

e1$roi<-factor(e1$roi,levels=c("CL","BEI","RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))

e1$ID<-factor(paste("e1_",e1$ID,sep=""))
colnames(e1)[3]<-"subject"
e1$expt<-"e1"
e1<-e1[,c(3,5,1,10,11,12)]
#write.table(e1,file="e1clean.txt")

e1q<-read.table("../data/Expt1.txt",header=TRUE)
e1q<-subset(e1q,region=="?")
colnames(e1q)[1]<-"subject"
e1q$subject<-factor(paste("e1_",e1q$subject,sep=""))
e1q<-e1q[,c(1,3,4,9,10)]
e1q$expt<-"e1"

head(e1q)
e1q<-e1q[,c(1,2,3,4)]

#dim(e1)
e1merged<-merge(e1,e1q,by.x=c("subject","item","condition"),
      by.y=c("subject","item","condition"))
#dim(e1merged)
#head(e1merged)

#head(e1merged)
e1merged<-e1merged[,c(1,2,3,4,7,5,6)]

write.table(e1merged,file="e1clean.txt")

## e2
e2<-read.csv("../data/Expt2.csv",header=TRUE)
e2$sub<-factor(paste("e2_",e2$sub,sep=""))
e2$item <- as.factor(e2$item)
e2$pos <- as.factor(e2$pos)
e2$posW <- as.character(e2$posW)
## relabel pos coding: 
#"RCSubj","RCV","de","headnoun","Adv","MainV","MainO"
## 2 embedded N RCSubj
## 3 rc verb RCV
## 4 de de
## 5 head noun headnoun
## 6 head noun+1: adv Adv
## 7 main verb MainV
## 8 main obj MainO
e2$roi<-ifelse(e2$relabelpos==2,"RCSubj",ifelse(e2$relabelpos==3,"RCV",
ifelse(e2$relabelpos==4,"de",
ifelse(e2$relabelpos==5,"headnoun",
ifelse(e2$relabelpos==6,"Adv",
ifelse(e2$relabelpos==7,"MainV",
ifelse(e2$relabelpos==8,"MainO",NA)))))))

head(e2)
colnames(e2)[1]<-"subject"
colnames(e2)[4]<-"condition"
e2<-e2[,c(1,3,4,8,9,13)]
head(e2)

e2$expt<-"e2"

colnames(e2)[5]<-"correct"

write.table(e2,file="e2clean.txt")
```


```{r loade1}
e1<-read.table("../data/e1clean.txt",header=TRUE)
```

```{r analysise1}
e1$CL<-ifelse(e1$condition%in%c("a","c"),-1,1)
e1$BEI<-ifelse(e1$condition%in%c("a","b"),-1,1)
e1$CLxBEI<-ifelse(e1$condition%in%c("b","c"),-1,1)

## nested contrasts:
## condition a: no cl no bei
## condition b: cl, no bei (regions are moved forward by 1) cl in region 0
## condition c: no cl, bei (regions are moved forward by 1) bei in region 0
## condition d: cl, be (regions are moved forward by 1), cl in region 0, bei in region 1

#       a    b     c     d
#Bei1   -1   0     1     0
#Bei2   0    -1    0     1
#CL    -1   1    -1      1

e1$BEI_inNoCL <- ifelse(e1$condition=="a",-1,
                     ifelse(e1$condition=="c",1,0))
e1$BEI_inCL <- ifelse(e1$condition=="b",-1,
                  ifelse(e1$condition=="d",1,0))

## 
#boxplot(-1000/RT~condition,e1)
#boxcox(e1$RT~e1$condition*e1$subject)

e1$rrt<- -1000/e1$RT

#xtabs(~CL+condition,e1)
#xtabs(~BEI+condition,e1)
#xtabs(~CLxBEI+condition,e1)

#xtabs(~CL+item,e1)
#xtabs(~BEI+ID,e1)
#tabs(~CLxBEI+item,e1)

## RCSubj
m1<-lmer(rrt~CL+BEI+CLxBEI+
           (1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1,roi=="RCSubj"))
#summary(m1)

##RCV
m2<-lmer(rrt~CL+BEI+CLxBEI+(1|subject)+
           (1|item),subset(e1,roi=="RCV"))

#summary(m2)

library(car)
qqPlot(residuals(m2))

##de main effects and interactions
m3<-lmer(rrt~CL+BEI+CLxBEI+(1+BEI+CLxBEI||subject)+
           (1+CL+BEI||item),
         subset(e1,roi=="de"))
summary(m3)

## nested comparison:
#xtabs(~BEI_inNoCL + condition, subset(e1,roi=="de"))
#xtabs(~BEI_inCL + condition, subset(e1,roi=="de"))
#xtabs(~BEI + condition, subset(e1,roi=="de"))

m3nested<-lmer(rrt~BEI_inNoCL + BEI_inCL + CL + (1+BEI_inNoCL + BEI_inCL ||subject)+
                 (CL||item),
               subset(e1,roi=="de"))

round(with(subset(e1,roi=="de"),tapply(RT,condition,mean)))

summary(m3nested)

##headnoun
m4<-lmer(rrt~CL+BEI+CLxBEI+(1+CLxBEI||subject)+
           (1|item),
         subset(e1,roi=="headnoun"))
summary(m4)

qqPlot(residuals(m4))

##Adv main effects
m5<-lmer(rrt~CL+BEI+CLxBEI+(1|subject)+
           (1+CL+BEI||item),
         subset(e1,roi=="Adv"))
summary(m5)

qqPlot(residuals(m5))

##MainV main effects
m6<-lmer(rrt~CL+BEI+CLxBEI+
           (1|subject)+
           (1+BEI+CLxBEI||item),
         subset(e1,roi=="MainV"))
summary(m6)

qqPlot(residuals(m6))

##MainO main effects CL effect
m7<-lmer(rrt~CL+BEI+CLxBEI+(1+BEI+CLxBEI||subject)+
           (1|item),
         subset(e1,roi=="MainO"))
summary(m7)
```

```{r}
resultse1<-rbind(coeflmer(m1), coeflmer(m2), coeflmer(m3),
      coeflmer(m4), coeflmer(m5), coeflmer(m6), coeflmer(m7))
contrasts<-rownames(resultse1)
contrasts<-factor(contrasts,levels=c("CL","BEI","CLxBEI"))

resultse1<-data.frame(resultse1)
colnames(resultse1)<-c("estimate","se","t")
resultse1<-cbind(contrasts,resultse1)

rois<-rep(c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"),each=3)
rois<-factor(rois,levels=c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))
resultse1<-cbind(resultse1,rois)

resultse1$lower<-resultse1$estimate-2*resultse1$se
resultse1$upper<-resultse1$estimate+2*resultse1$se
resultse1$expt<-"e1"
```

Statistical significant at threshold 2.4:

```{r}
resultse1[which(abs(resultse1$t) >= 2.4),c(5,1,2,3,4)]
```

Plot of results:

```{r}
plotresults(resultse1,maintitle="Experiment 1")
```

Appendix means and SEs in milliseconds:


```{r}
e1crit<-subset(e1,roi%in%c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))

colnames(e1crit)[4]<-"rt"

e1crit$roi<-as.factor(as.character(e1crit$roi))
#unique(e1crit$roi)

e1crit$roi<-factor(e1crit$roi,levels=c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))

M.id.w<-regionmeans(e1crit)
```

```{r results="asis"}
library(xtable)
print(xtable(M.id.w[,c(1,2,6,7,3,4)],digits=c(0,0,0,0,0,0,0)),
      type="html")
```

Rename conditions for plot:

```{r fig.width=8,fig.height=8}
M.id.w$condition<-factor(paste(M.id.w$condition,M.id.w$classifier,M.id.w$bei,sep=" "))

levels(M.id.w$condition)<-c("a NoCL NoBEI", "b CL NoBEI", "c NoCL BEI", "d CL BEI")

regionplot(M.id.w,maintitle="Experiment 1")
```


```{r fig.width=8,fig.height=8}
multiplot(plotresults(resultse1,maintitle="Experiment 1"),regionplot(M.id.w,maintitle="Experiment 1"),cols=1)
```

# Experiment 2

```{r loade2}
e2<-read.table("../data/e2clean.txt",header=TRUE)
```

```{r analysise2}
e2$CL<-ifelse(e2$condition%in%c("a","c"),-1,1)
e2$BEI<-ifelse(e2$condition%in%c("a","b"),-1,1)
e2$CLxBEI<-ifelse(e2$condition%in%c("b","c"),-1,1)

## nested contrasts:
## condition a: no cl no bei
## condition b: cl, no bei (regions are moved forward by 1) cl in region 0
## condition c: no cl, bei (regions are moved forward by 1) bei in region 0
## condition d: cl, be (regions are moved forward by 1), cl in region 0, bei in region 1

#       a    b     c     d
#Bei1   -1   0     1     0
#Bei2   0    -1    0     1
#CL    -1   1    -1      1

e2$BEI_inNoCL <- ifelse(e2$condition=="a",-1,
                     ifelse(e2$condition=="c",1,0))
e2$BEI_inCL <- ifelse(e2$condition=="b",-1,
                  ifelse(e2$condition=="d",1,0))

#dim(e2)

e2$rrt<- -1000/e2$rt

#summary(e2$rt)

#dim(e2)
e2<-subset(e2,rt>150)
#dim(e2)
#100*(11136-11128)/11136
## .07% of the data removed

## RCSubj
m1<-lmer(rrt~CL+BEI+CLxBEI+
           (1+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e2,roi=="RCSubj"))
summary(m1)

#qqPlot(residuals(m1))

##RCV
m2<-lmer(rrt~CL+BEI+CLxBEI+(1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),subset(e2,roi=="RCV" & rt>150))
summary(m2)

qqPlot(residuals(m2))

##de main effects and interactions
summary(lmer(rrt~CL+BEI+CLxBEI+(1+CL+BEI+CLxBEI||subject)+
       (1+CL+BEI+CLxBEI||item),subset(e2,roi=="de")))

m3<-lmer(rrt~CL+BEI+CLxBEI+(1+CLxBEI||subject)+
           (1+CLxBEI||item),
         subset(e2,roi=="de"))
summary(m3)

## nested comparison:
#xtabs(~BEI_inNoCL + condition, subset(e1,roi=="de"))
#xtabs(~BEI_inCL + condition, subset(e1,roi=="de"))
#xtabs(~BEI + condition, subset(e1,roi=="de"))

m3nested<-lmer(rrt~BEI_inNoCL + BEI_inCL + CL + 
                   (1 + BEI_inCL ||subject)+
                 (1 |item),
               subset(e2,roi=="de"))

summary(m3nested)

##headnoun
m4<-lmer(rrt~CL+BEI+CLxBEI+(1+CL+CLxBEI||subject)+
           (1+BEI||item),
         subset(e2,roi=="headnoun"))
summary(m4)

qqPlot(residuals(m4))

##Adv main effects
m5<-lmer(rrt~CL+BEI+CLxBEI+(1+BEI+CLxBEI||subject)+
           (1+CL||item),
         subset(e2,roi=="Adv"))
summary(m5)

#qqPlot(residuals(m5))

##MainV main effects
m6<-lmer(rrt~CL+BEI+CLxBEI+
           (1+BEI+CLxBEI||subject)+
           (1+CL+BEI||item),
         subset(e2,roi=="MainV"))
summary(m6)

#qqPlot(residuals(m6))

##MainO main effects CL effect
m7<-lmer(rrt~CL+BEI+CLxBEI+(1+CLxBEI||subject)+
           (1+CL||item),
         subset(e2,roi=="MainO"))
summary(m7)

qqPlot(residuals(m7))
```

Plot results:

```{r}
resultse2<-rbind(coeflmer(m1), coeflmer(m2), coeflmer(m3),
      coeflmer(m4), coeflmer(m5), coeflmer(m6), coeflmer(m7))
contrasts<-rownames(resultse2)
contrasts<-factor(contrasts,levels=c("CL","BEI","CLxBEI"))

resultse2<-data.frame(resultse2)
colnames(resultse2)<-c("estimate","se","t")
resultse2<-cbind(contrasts,resultse2)

rois<-rep(c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"),each=3)
rois<-factor(rois,levels=c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))
resultse2<-cbind(resultse2,rois)

resultse2$lower<-resultse2$estimate-2*resultse2$se
resultse2$upper<-resultse2$estimate+2*resultse2$se
resultse2$expt<-"e2"
```

```{r}
## comparing results:
subset(resultse1,abs(t)>2.4)[,c(5,1,2,3,4)]
subset(resultse2,abs(t)>2.4)[,c(5,1,2,3,4)]
```

```{r fig.width=8,fig.height=8}
plotresults(resultse2,maintitle="Experiment 2")
```

```{r}
e2crit<-subset(e2,
               roi%in%c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))

e2crit$roi<-as.factor(as.character(e2crit$roi))
unique(e2crit$roi)

e2crit$roi<-factor(e2crit$roi,
                   levels=c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))

M.id.w<-regionmeans(e2crit)
#M.id.w[,c(1,2,6,7,3,4)]
```

```{r fig.width=8,fig.height=8}
M.id.w$condition<-factor(paste(M.id.w$condition,M.id.w$classifier,M.id.w$bei,sep=" "))

levels(M.id.w$condition)<-c("a NoCL NoBEI", "b CL NoBEI", "c NoCL BEI", "d CL BEI")

regionplot(M.id.w,maintitle="Experiment 2",legendpos=c(0,0.4))
```

```{r results="asis"}
print(xtable(M.id.w[,c(1,2,6,7,3,4)],digits=c(0,0,0,0,0,0,0)),
      type="html")
```

```{r fig.width=8,fig.height=8}
multiplot(plotresults(resultse2,maintitle="Experiment 2"),
          regionplot(M.id.w,maintitle="Experiment 2",legendpos=c(0,0.4)),ncol=1)
```

```{r plotresultse1e2,fig.width=8,fig.height=8}
multiplot(plotresults(resultse1,maintitle="Experiment 1"),
          plotresults(resultse2,maintitle="Experiment 2"),ncol=1)
```

# Results

Since we analyzed seven regions of interest that are correlated, a conservative approach is to correct the $\alpha$ for multiple corrections. We use the Sidak correction; nothing changes if we use the Bonferroni.

```{r}
sidak_alpha<-1-(1-0.05)^(1/7)
(crit_t_sidak<-round(abs(qnorm(sidak_alpha)),1))
bonf_alpha<-0.05/7
(crit_t_bonf<-round(abs(qnorm(bonf_alpha)),1))
```


## With Sidak correction

We clearly see that the effects are stronger in Experiment 2:

```{r printresults,echo=FALSE,eval=TRUE,cache=FALSE,results="asis"}
print(xtable(subset(resultse1,abs(t)>crit_t_sidak)),type="html")
print(xtable(subset(resultse2,abs(t)>crit_t_sidak)),type="html")
```

# Combined analysis:

```{r combinedanalysis}

colnames(e1)
colnames(e1)[4]<-"rt"
colnames(e2)

e1e2<-rbind(e1,e2)

e1e2$expt<-ifelse(e1e2$expt=="e1",-1,1)

## RCSubj
m1<-lmer(rrt~(CL+BEI+CLxBEI)*expt+
           (1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1e2,roi=="RCSubj"))
summary(m1)

qqPlot(residuals(m1))

##RCV
m2<-lmer(rrt~(CL+BEI+CLxBEI)*expt+(1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),subset(e1e2,roi=="RCV"))
summary(m2)

qqPlot(residuals(m2))

##de main effects and interactions
m3<-lmer(rrt~(CL+BEI+CLxBEI)*expt+(1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1e2,roi=="de"))
summary(m3)

##headnoun
m4<-lmer(rrt~(CL+BEI+CLxBEI)*expt+(1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1e2,roi=="headnoun"))
summary(m4)

qqPlot(residuals(m4))

##Adv main effects
m5<-lmer(rrt~(CL+BEI+CLxBEI)*expt+(1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1e2,roi=="Adv"))
summary(m5)

qqPlot(residuals(m5))

##MainV main effects
m6<-lmer(rrt~(CL+BEI+CLxBEI)*expt+
           (1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1e2,roi=="MainV"))
summary(m6)

#qqPlot(residuals(m6))

##MainO main effects CL effect
m7<-lmer(rrt~(CL+BEI+CLxBEI)*expt+(1+CL+BEI+CLxBEI||subject)+
           (1+CL+BEI+CLxBEI||item),
         subset(e1e2,roi=="MainO"))
summary(m7)
```

```{r}
resultse1e2<-rbind(coeflmer(m1), coeflmer(m2), coeflmer(m3),
      coeflmer(m4), coeflmer(m5), coeflmer(m6), coeflmer(m7))
contrasts<-rownames(resultse1e2)
contrasts<-factor(contrasts,levels=c("CL","BEI","CLxBEI"))

resultse1e2<-data.frame(resultse1e2)
colnames(resultse1e2)<-c("estimate","se","t")
resultse1e2<-cbind(contrasts,resultse1e2)

rois<-rep(c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"),each=3)
rois<-factor(rois,levels=c("RCSubj","RCV","de","headnoun","Adv","MainV","MainO"))
resultse1e2<-cbind(resultse1e2,rois)

resultse1e2$lower<-resultse1e2$estimate-2*resultse1e2$se
resultse1e2$upper<-resultse1e2$estimate+2*resultse1e2$se
resultse1e2$expt<-"e1e2"
```

```{r plotallresults,fig.width=8,fig.height=8}
multiplot(plotresults(resultse1,maintitle="Experiment 1"),
          plotresults(resultse2,maintitle="Experiment 2"),
          plotresults(resultse1e2,maintitle="Experiments 1 and 2 combined"),ncol=1)

```

```{r processRmd, echo=FALSE, eval=FALSE}
## not run:
rmarkdown::render('./WuEtAlCogSci2016-vignette.Rmd')
```

