---
title: "Testing multiple W parameter settings and model checking"
author: "Paul Mätzig"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{low W parameter analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro 

This follows the same workflow as the demo analysis vignette provided with this
package. However, instead of two datasets with only a slight variation in the
$W$ parameter in ACT-R, here we will see five different, low $W$ parameter
settings. For each of those settings, the simulation was repeated four times in
order to enable us to do quick model checking in the end.

## Preparing the data

```{r}
library(ACTRPredictions2016)
library(dplyr)
library(ggplot2)
```

```{r}
p <- get_parameters("../data/GG-EXP1-2016112-1759/paramsearch.txt")
head(p)
```

```{r}
fixation_list <- get_data("../data/GG-EXP1-2016112-1759/", p)
summary(fixation_list)
```

```{r}
data_list <- calculate_reading_measures(fixation_list)
summary(data_list)
head(data_list[[1]])
```

```{r}
data_list <- extract_main_emb_V(data_list)
head(data_list[[1]])
```

```{r}
results_list <- mean_se_by_region(data_list, measure="TFT")
summary(results_list)
print(head(results_list[[1]]))
print(results_list[[1]]$`mean(TFT)`)
print(results_list[[1]]$`standard_error(TFT)`)
```

```{r, fig_width=15}
cowplot::plot_grid(plot_by_region(results_list[[1]], maintitle="W 0.2 rep 1", legendpos=c(0, 0.2)),
                   plot_by_region(results_list[[6]], maintitle="W 0.2 rep 2", legendpos=c(0, 0.2)),
                   plot_by_region(results_list[[11]], maintitle="W 0.2 rep 3", legendpos=c(0, 0.2)),
                   plot_by_region(results_list[[16]], maintitle="W 0.2 rep 4", legendpos=c(0, 0.2)),
                   labels=c("(a)", "(b)", "(c)", "(d)") )
```
